---
title: "edld_650_dare_01"
author: "Dodjivi Amekoudji & Michelle Cui"
format: html
editor: visual
---

EDLD 650 Advanced Seminar in Educational Research Methods\
Professor: David D. Liebowitz\
Apr 15, 2026

## Data Analysis and Replication Exercise (DARE) 01

### Introduction

You will replicate and extend analysis from Liebowitz, Porter and Bragg. In the file EDLD 650 DARE 1.csv, you will find the variables listed in Table 1. These data have been aggregated to the state level and altered to preserve confidentiality; thus your results will diﬀer (slightly) from those in the paper. Please make sure that your graph axes are labeled, the variables in your tables have comprehensible labels, and your table notes are complete.\

```{r}
# Load package
library(pacman)
p_load(here, tidyverse, lme4, haven, fixest)
```

```{r}
# Load the data
df <- read_csv(here("data/EDLD_650_DARE_1.csv")) 
str(df)

```

#### A. Data Management Tasks

A1. Convert the raw counts of enrollment by race/ethnicity into percent- ages (i.e., divide the enrollment count for each ethno-racial category by total enrollment). For programming eﬃciency, can you use a function to do this task?

```{r}
df <- df %>% 
  mutate(enroll_other_percent = enroll_OTHER/enroll,
          enroll_AM_percent = enroll_AM/enroll,
          enroll_ASIAN_percent = enroll_ASIAN/enroll,
          enroll_HISP_percent = enroll_HISP/enroll,
          enroll_BLACK_percent = enroll_BLACK/enroll,
          enroll_WHITE_percent = enroll_WHITE/enroll)

summary(df$enroll_AM_percent)
```

A2. Generate dichotomous policy predictor variables that take the value of 1 in state-year observations in which the policy is in place. Call them *eval*, *class_remove* and *suspension*. They should take the value of 0 in years during which these policies were not in place. Also, generate a running time variable (*run_time*) that reflects how far or close the state-year observation is from the implementation of higher- stakes teacher evaluation and a variable that permits the eﬀects of the evaluation policy to vary (linearly) over time (evalXyear). How will you deal with states that never implement evaluation? Do that too.

```{r}
df <- df %>% 
  mutate(
    eval = case_when(!is.na(eval_year) & school_year >= eval_year ~ 1, TRUE ~ 0),
    class_remove = case_when(!is.na(class_remove_year) & school_year >= class_remove_year ~ 1,
                             TRUE ~ 0),
         suspension = case_when(!is.na(suspension_year) & school_year >= suspension_year ~ 1, 
                                TRUE ~ 0),
         run_time = case_when(is.na(eval_year) ~ 0, 
                              TRUE ~ as.numeric(school_year - eval_year)),
         evalXyear = eval*run_time
    
  )
```
